Replacing standard galaxy profiles with \\ mixtures of Gaussians

### 2012-10-23 documents/mixtures/mixture_models.tex

% to-do
% -----
% - HOGG: go through remining github issues and close or fix
% - HOGG: send to Bertin and Rix and Peng and Lupton and Bundy for comments
% - HOGG: submit!
% - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
% - HOGG: make what I say about M=6 lux and M=8 lux true -- ie, edit the Tractor code!
% - DSTN: can we be more specific about how Bosch is using them? I would like to cite users.
% - HOGG: should we make Machine Readable Tables for PASP?  Yes.

@@

We need good performance in two-dimensional image synthesis (for
fitting) because, with \thetractor\ (Lang \etal, forthcoming), we are
building a comprehensive model of all the imaging data we have; this
will have models of many millions of galaxies in imaging that contains
on the order of $10^{13}$ pixels.  We are basing our models on the
\project{Sloan Digital Sky Survey} (\sdss) Catalog galaxy models,
which only include exponential, de~Vaucouleurs, and composite (mixture
of the two) radial profiles.  In detail, in fact, the \sdss\ Catalog
models are small modifications of these; details
below.  So our goal here is to provide replacements for these models
in order to improve the performance of galaxy image modeling and
analysis software.  Some of the models we present have been used
previously in tools for precise photometry (\citealt{bundy}), and are
being used in the Large Synoptic Survey Telescope (LSST) analysis
codes along with shapelet representations for the PSF
(\citealt{bosch}).  These models could also be used to speed other
image-fitting systems, like the very successful \project{GALFIT}
(\citealt{galfit}).

@@

\begin{thebibliography}{70}
\bibitem[Bendinelli(1991)]{bendinelli}
Bendinelli,~O., 1991, \apj, 366, 599
\bibitem[Bendinelli \& Parmeggiani(1995)]{bendinelli2}
Bendinelli,~O. \& Parmeggiani,~G., 1995, \aj, 109, 572
\bibitem[Bosch(2010)]{bosch}
Bosch,~J., 2010, \aj, 140, 870
\bibitem[Bovy \etal(2011a)]{xd}
Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011, Ann. Appl. Stat., 5, 1657
\bibitem[Bovy \etal(2011b)]{xdqso}
Bovy,~J., \etal, 2011, \apj, 729, 141
\bibitem[Bovy \etal(2012)]{xdqsoz}
Bovy,~J. \etal, 2012, \aj, 749, 41
\bibitem[Bundy \etal(2012)]{bundy}
Bundy,~K. \etal, \aj, in press
\bibitem[Cappellari(2002)]{cappellari}
Cappellari,~M., 2002, \mnras, 333, 400
\bibitem[Ciotti \& Bertin(1999)]{ciotti}
Ciotti,~L. \& Bertin,~G., 1999, \aa, 352, 447
\bibitem[de~Vaucouleurs(1948)]{dev}
de~Vaucouleurs,~G., 1948, Annales d'Astrophysique, 11, 247
\bibitem[Emsellem \etal(1994)]{emsellem}
Emsellem,~E., Monnet,~G., Bacon,~R., \& Nieto,~J.-L., 1994, \aap, 285, 739
\bibitem[Lupton \etal(2001)]{lupton}
Lupton,~R., Gunn,~J.~E., Ivezic,~Z., Knapp,~G.~R., Kent,~S.~M., \& Yasuda,~N., 2001, ASPC, 238, 269
\bibitem[Peng \etal(2002)]{galfit}
Peng,~C.~Y., Ho, L.~C., Impey,~C.~D., \& Rix,~H.-W., 2002, \aj, 124, 266
\bibitem[S{\'e}rsic(1963)]{ser}
S{\'e}rsic,~J.~L., 1963, Boletin de la Asociacion Argentina de Astronomia La Plata Argentina, 6, 41
\end{thebibliography}

>>>>

% to-do
% -----
% - HOGG: go through remining github issues and close or fix
% - HOGG: send to Bertin and Rix and Peng and Lupton and Bundy for comments
% - HOGG: submit!
% - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
% - HOGG: make what I say about M=6 lux and M=8 lux true -- ie, edit the Tractor code!
% - HOGG: should we make Machine Readable Tables for PASP?  Yes.

@@

We need good performance in two-dimensional image synthesis (for
fitting) because, with \thetractor\ (Lang \etal, forthcoming), we are
building a comprehensive model of all the imaging data we have; this
will have models of many millions of galaxies in imaging that contains
on the order of $10^{13}$ pixels.  We are basing our models on the
\project{Sloan Digital Sky Survey} (\sdss) Catalog galaxy models,
which only include exponential, de~Vaucouleurs, and composite (mixture
of the two) radial profiles.  In detail, in fact, the \sdss\ Catalog
models are small modifications of these; details
below.  So our goal here is to provide replacements for these models
in order to improve the performance of galaxy image modeling and
analysis software.  Some of the models we present have been used
previously in tools for precise photometry (\citealt{bundy}), and are
being used in the Large Synoptic Survey Telescope (LSST) prototype
galaxy photometry pipeline (\citealt{shaw}), in which these models are
convolved with a shapelet representation of the PSF; because shapelets
are simply polynomial perturbations on a Gaussian, they can also be
convolved analytically with mixture-of-Gaussian galaxy models
(\citealt{bosch}).  These models could also be used to speed other
image-fitting systems, like the very successful \project{GALFIT}
(\citealt{galfit}).

@@

\begin{thebibliography}{70}
\bibitem[Bendinelli(1991)]{bendinelli}
Bendinelli,~O., 1991, \apj, 366, 599
\bibitem[Bendinelli \& Parmeggiani(1995)]{bendinelli2}
Bendinelli,~O. \& Parmeggiani,~G., 1995, \aj, 109, 572
\bibitem[Bosch(2010)]{bosch}
Bosch,~J., 2010, \aj, 140, 870
\bibitem[Bovy \etal(2011a)]{xd}
Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011, Ann. Appl. Stat., 5, 1657
\bibitem[Bovy \etal(2011b)]{xdqso}
Bovy,~J., \etal, 2011, \apj, 729, 141
\bibitem[Bovy \etal(2012)]{xdqsoz}
Bovy,~J. \etal, 2012, \aj, 749, 41
\bibitem[Bundy \etal(2012)]{bundy}
Bundy,~K. \etal, \aj, in press
\bibitem[Cappellari(2002)]{cappellari}
Cappellari,~M., 2002, \mnras, 333, 400
\bibitem[Ciotti \& Bertin(1999)]{ciotti}
Ciotti,~L. \& Bertin,~G., 1999, \aa, 352, 447
\bibitem[de~Vaucouleurs(1948)]{dev}
de~Vaucouleurs,~G., 1948, Annales d'Astrophysique, 11, 247
\bibitem[Emsellem \etal(1994)]{emsellem}
Emsellem,~E., Monnet,~G., Bacon,~R., \& Nieto,~J.-L., 1994, \aap, 285, 739
\bibitem[Lupton \etal(2001)]{lupton}
Lupton,~R., Gunn,~J.~E., Ivezic,~Z., Knapp,~G.~R., Kent,~S.~M., \& Yasuda,~N., 2001, ASPC, 238, 269
\bibitem[Peng \etal(2002)]{galfit}
Peng,~C.~Y., Ho, L.~C., Impey,~C.~D., \& Rix,~H.-W., 2002, \aj, 124, 266
\bibitem[S{\'e}rsic(1963)]{ser}
S{\'e}rsic,~J.~L., 1963, Boletin de la Asociacion Argentina de Astronomia La Plata Argentina, 6, 41
\bibitem[Shaw(2012)]{shaw}
Shaw,~R.~A., ed., 2012, LSST Data Challenge Handbook (Version 2.0; Tucson, AZ: LSST Corp.)
\end{thebibliography}
